Scaffolder - Software for Microbial Genome Scaffolding.

### 2011-01-25 text/src/article.tex
Fix broken header
Results section for scaffolder file format

Individual reads produced from nucleotide sequencing can be assembled into
contiguous regions (``contigs'') of many overlapping individual reads. These
reads may however not represent a complete genome sequence where all contigs do
not result in a single complete sequence. The software outlined here
``Scaffolder'' allows these contigs to be further joined together into
a continuous single sequence. This selection, order and direction of contigs
into a final draft genome is often described as the `scaffold'. Scaffolder
allows the manual or computational specification of a genome scaffold by
specifying which contig sequences are used in simple to edit configuration
files.

The scaffold file specifies how contigs should be joined together in the draft
genome sequence. The raw nucleotide sequence for each contig is stored in
a fasta file, as might be produced from assembly of sequencing reads using
programs such as Newbler\cite{miller2010} or ABySS\cite{simpson2009}.
Scaffolder then the contig to be selected by specifying the name of the contig
in the scaffold file. The scaffold file also allows a sub-sequence or the
reverse complement of a contig to used. This therefore preserves the original
sequence data but also allowing sequences to manipulated in the final genome
build.

The majority of a genome can be estimated from scaffolding assembled contigs
into the most parsimonious build. There may however still be internal gaps in
the contigs from mate pair assembly based on estimated mate pair insert size.
Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
and filled - a common process in finishing a genome sequence. Inserts may be
treated the same way as a contig where the raw sequence, as may be produced by
PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
in the contig.

Scaffolder allows the specification of unknown regions in the scaffold. These
are areas of the genome where the nucleotide sequence is unknown but the
approximate size region may be estimated from mate pair insert size or from
mapping contigs to reference genome and estimating the size of the gap. These
unknown regions allow a continuous draft sequence to be scaffolded but whilst
still highlighting regions waiting to be resolved.

@@

The genome scaffold is defined YAML format. YAML is an easy to edit data format
that is both machine readable and easy to edit manually.

INSERT FIGURE WITH EXAMPLE LAYOUT OF SCAFFOLD

  * Screen shot with example file
  * Illustrate three types of region in the scaffold file
  * Explanation of what each region does
  * Expected output from each region

Scaffolder provides a ruby API that allows the scaffold and designated
sequences to be manipulated through a ruby programme. This allows scaffolder to
be integrated into exiting genomics works flows such as using Rake. Scaffolder
all provides a command line binary for the manipulation of the scaffold. This
binary behaves as a standard Unix command line interface returning appropriate
exit statuses and providing manual pages for each of the commands. These two
methods are outlined in the following two sections.

\subsection*{Command line usage} %{{{2

@@

\subsection*{Programmatic usage of API} %{{{2

@@

\subsection*{Installation} %{{{2

  * Install through RubyGems
  * Make sure the bin is in the class path

\section*{Conclusions} %{{{1
  Text for this section \ldots
% }}}
% {{{ Availability and Requirements
\section*{Availability and Requirements}

@@

% }}}
% {{{ Author's Information
\section*{Authors contributions}
    Text for this section \ldots
% }}}
% {{{ Acknowledgements
\section*{Acknowledgements}
  \ifthenelse{\boolean{publ}}{\small}{}
  Text for this section \ldots
%}}}
% {{{ Bibliography
{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
 \bibliographystyle{bmc_article}  % Style BST file
  \bibliography{article} }     % Bibliography file (usually '*.bib' ) 
%}}}
% {{{ Figures
\ifthenelse{\boolean{publ}}{\end{multicols}}{}

\section*{Figures}
  \subsection*{Figure 1 - Sample figure title}
      A short description of the figure content
      should go here.

  \subsection*{Figure 2 - Sample figure title}
      Figure legend text.
% }}}
% {{{ Tables
\section*{Tables}
  \subsection*{Table 1 - Sample table title}
    Here is an example of a \emph{small} table in \LaTeX\ using  
    \verb|\tabular{...}|. This is where the description of the table 
    should go. \par \mbox{}
    \par
    \mbox{
      \begin{tabular}{|c|c|c|}
        \hline \multicolumn{3}{|c|}{My Table}\\ \hline
        A1 & B2  & C3 \\ \hline
        A2 & ... & .. \\ \hline
        A3 & ..  & .  \\ \hline
      \end{tabular}
      }
  \subsection*{Table 2 - Sample table title}
    Large tables are attached as separate files but should
    still be described here.
% }}}
% {{{ Additional Files
\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

@@

\end{bmcformat}
\end{document}
% }}}

>>>>

Individual reads produced from nucleotide sequencing can be assembled into
contiguous regions (``contigs'') of many overlapping individual reads. These
reads may however not represent a complete genome sequence where all contigs do
not result in a single complete sequence. The software outlined here
``Scaffolder'' allows these contigs to be further joined together into
a continuous single sequence. This selection, order and direction of contigs
into a final draft genome is often described as the `scaffold'. Scaffolder
allows the manual or computational specification of a genome scaffold by
specifying which contig sequences are used in simple to edit configuration
files. \pb

The scaffold file specifies how contigs should be joined together in the draft
genome sequence. The raw nucleotide sequence for each contig is stored in
a fasta file, as might be produced from assembly of sequencing reads using
programs such as Newbler\cite{miller2010} or ABySS\cite{simpson2009}.
Scaffolder then the contig to be selected by specifying the name of the contig
in the scaffold file. The scaffold file also allows a sub-sequence or the
reverse complement of a contig to used. This therefore preserves the original
sequence data but also allowing sequences to manipulated in the final genome
build. \pb

The majority of a genome can be estimated from scaffolding assembled contigs
into the most parsimonious build. There may however still be internal gaps in
the contigs from mate pair assembly based on estimated mate pair insert size.
Scaffolder allows inserts to be added to scaffolded so that gaps may be bridged
and filled - a common process in finishing a genome sequence. Inserts may be
treated the same way as a contig where the raw sequence, as may be produced by
PCR is stored in a fasta file, and then trimmed and/or reversed to fill the gap
in the contig. \pb

Scaffolder allows the specification of unknown regions in the scaffold. These
are areas of the genome where the nucleotide sequence is unknown but the
approximate size region may be estimated from mate pair insert size or from
mapping contigs to reference genome and estimating the size of the gap. These
unknown regions allow a continuous draft sequence to be scaffolded but whilst
still highlighting regions waiting to be resolved. \pb

@@

The genome scaffold is wriiten in YAML format. YAML is a data format designed
to be both machine readable and easy to manually edit. An example of this
format illustrating all the features of scaffolder is shown in Figure 1. \pb

\subsubsection*{Simple sequence region} %{{{3

The first line of the file begins with three dashes. This is required to
signify the start of a YAML document. Lines 2 begins with a dash character `-'
which denotes the start of an entry in the sequence scaffold, in this case the
first entry. Line 3 is indented by two spaces which therefore means this line
contains an entry type. The ``sequence'' tag is used indicate a sequence is
inserted at this region in the scaffold. The following line contain is
indented by two spaces and indicates that this an attribute of the sequence
entry. In this case the ``source'' keyword is used to indicate used which
sequence should be inserted into the scaffold. This sequence is indentified by
the first word of the fasta header of the selected sequence. This indicates
a simple sequence region added to the scaffold. \pb

\subsubsection*{Unresolved seqeunce region} %{{{3

The following line has no indentation and begins with dash indicating the next
entry in the scaffold. The following entry is indentified by the ``unresolved''
tag. This identifies a region of unknown nucleotide sequence but is included
because the approximate size of the region is known. The following line defines
that the size of this unknown region in the genome scaffold should be 20
nucleotides. This therefore results in the region being filled with 20 `N'
characters in draft sequence. \pb

\subsubsection*{Trimmed sequence region with multiple inserts} %{{{3

The following entry is another sequence using the sequence from 'sequence2' in
the scaffold file. This sequence entry has three further attributes in addition
to defining the source sequence. The first two are start and stop. These are
used to trim the final sequence inserted into the scaffold. In this case the
first two nucleotides of the sequence are removed so that that sequence starts
at position 3 of the original raw sequence. Similarly the sequence after
position is trimmed away after position 100. These sequence is also reverse
complemented using the ``reverse'' tag. Reverse complementation is performed
after the sequence has been trimmed to the specified coordinates. \pb

The final attribute for this sequence entry defines inserts to be added to the
sequence. These are additional regions of sequence that can be added to
a sequence entry to alter the existing sequence such to fill gaps or correct
sequencing errors. Inserts are defined using the ``inserts'' tag. Each insert
is then listed similarly to the entries in the scaffold: beginning with a dash
on one line then indented by an additional two spaces on the following lines.
\pb

Inserts are listed similarly to sequence entries where reverse, start and stop
attributes can be used to trim and reverse the sequence. The `source' tag is
used to indentify which fasta sequence entry should be used as the insert.  The
insert has two additional tags ``open'' and ``close'' which determine where the
insert is added to the hosting sequence. These coordinates are inclusive
replaceing the coordinate nucleotides. The open and close positions need not be
the same size as the insert size. Any difference in size to the replaced
nucleotides will result in the length on the encoding sequence being adjusted.
\pb

The next insert entry shows an insert with only the open position specified.
This is to illustrate that only the one of the `open' or `close' keywords is
sufficient to add an insert. If only one of these words is used then the
opposing insert position is calculated from the length on the insert sequence
in the fasta file. This feature is designed for using inserts to close gaps
where the length of the insert may not completely fill the gap but can still be
used to fill a portion of the gap sequence. \pb

\subsubsection*{Scaffolder file attribute processing} %{{{3

Each scaffolder sequence entry can have multiple attributes which applied in
different order can result in a different sequence. For instance reverse
complementing the sequence then adding inserts will have a different effect
compared with performing these actions in the reverse order. Scaffolder
therefore follows the following protocol for processing each entry in the
scaffold file.

\paragraph{Process inserts}

\subsection*{Scaffolder usage} %{{{2

Scaffolder provides a ruby API that allows the scaffold and designated
sequences to be manipulated through a ruby programme. This allows scaffolder to
be integrated into exiting genomics works flows such as using Rake.  Scaffolder
all provides a command line binary for the manipulation of the scaffold. This
binary behaves as a standard Unix command line interface returning appropriate
exit statuses and providing manual pages for each of the commands. These two
methods are outlined in the following two sections. \pb

\subsubsection*{Installation} %{{{3

  * Install through RubyGems

  * Make sure the bin is in the class path

\subsubsection*{Command line interface} %{{{3

@@

\subsubsection*{Accessible programmatic interface} %{{{3

@@

\subsection*{Discussion of scaffolder approach} %{{{2

\section*{Conclusions} %{{{1

Text for this section \ldots }}} {{{ Availability and Requirements

\section*{Availability and Requirements}

@@

% }}}
% {{{ Author's Information
\section*{Authors contributions}
    Text for this section \ldots
% }}}
% {{{ Acknowledgements
\section*{Acknowledgements}
  \ifthenelse{\boolean{publ}}{\small}{}
  Text for this section \ldots
%}}} {{{1 Bibliography
{\ifthenelse{\boolean{publ}}{\footnotesize}{\small}
 \bibliographystyle{bmc_article}  % Style BST file
  \bibliography{article} }     % Bibliography file (usually '*.bib' )

\ifthenelse{\boolean{publ}}{\end{multicols}}{}

\section*{Figures} %{{{1

  \subsection*{Figure 1 - Sample figure title}

  A short description of the figure content should go here.

\section*{Tables} %{{{1

  \subsection*{Table 1 - Sample table title}
    Here is an example of a \emph{small} table in \LaTeX\ using  
    \verb|\tabular{...}|. This is where the description of the table 
    should go. \par \mbox{}
    \par
    \mbox{
      \begin{tabular}{|c|c|c|}
        \hline \multicolumn{3}{|c|}{My Table}\\ \hline
        A1 & B2  & C3 \\ \hline
        A2 & ... & .. \\ \hline
        A3 & ..  & .  \\ \hline
      \end{tabular}
      }
  \subsection*{Table 2 - Sample table title}
    Large tables are attached as separate files but should
    still be described here.

\section*{Additional Files} %{{{1

  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

@@

\end{bmcformat}

\end{document}
